---
title: "Glycine max Williams 82 v4 (Glycine_max_v2.1) Release 62 Annotation Package"
author: "Kabilan S"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Glycine max Williams 82 v4 Annotation Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This package delivers Ensembl‑based gene annotations for *Glycine max* (soybean)
assembly Williams 82 v4 (Glycine_max_v2.1), Release 62. Gene IDs follow the
**Wm82.a4.v1** nomenclature (`Glyma.XXGXXXXXX`). Chromosome lengths are included.

**Note:** Transcript and exon features are not available in this build because the
source GFF3 contained only gene‑level records. All gene‑level operations (TSS
retrieval, promoter‑region generation, filtering by gene ID or biotype) work as
expected.

# Installation

From the source tarball:

```{r installing, eval = FALSE}
install.packages("EnsDb.Gmax.Wm82.v62", repos = NULL, type = "source")
```

# Loading the package

```{r loading}
library(EnsDb.Gmax.Wm82.v62)
edb <- tryCatch(
  EnsDb.Gmax.Wm82.v62,
  error = function(e) {
    message("Database not available (maybe offline). Skipping example.")
    return(NULL)
  }
)
```

# Available columns and filters

```{r overview, eval = !is.null(edb)}
columns(edb)
supportedFilters(edb)
```

# Retrieve all genes

`genes()` returns a `GRanges` object with gene‑level information.

```{r genes, eval = !is.null(edb)}
gene_gr <- genes(edb)
head(gene_gr)
```

# Filter genes by ID

```{r filter_gene, eval = !is.null(edb)}
genes(edb, filter = GeneIdFilter("Glyma.01G000100"))
```

# Filter genes by biotype

```{r filter_gene_biotype, eval = !is.null(edb)}
genes(edb, filter = GeneBiotypeFilter("protein_coding"))
```

# Working with chromosome names

The database stores chromosome names as bare numbers (`1`, `2`, …). To add a
common prefix (e.g., `Gm01`), first restrict to the 20 nuclear chromosomes and
then modify the seqlevels:

```{r chromosomes, eval = !is.null(edb)}
gene_gr <- genes(edb)
library(GenomeInfoDb)
gene_gr_main <- keepStandardChromosomes(gene_gr, pruning.mode = "coarse")
seqlevels(gene_gr_main)   # bare numbers 1–20

# Add 'Gm' prefix
new_levels <- paste0("Gm", sprintf("%02d", as.integer(seqlevels(gene_gr_main))))
seqlevels(gene_gr_main) <- new_levels
seqlevels(gene_gr_main)
```

# Session information

```{r session}
sessionInfo()
```